---
title: 'ROSMAP snRNA-seq: MEGENA network analysis'
author: "Qian Wang, Minghui Wang, Won-Min Song"
date: "November 29, 2021"
output: html_document
---

## Example pipeline to perform MEGENA network inference
```{r setup environment, eval = TRUE, echo = TRUE, results = "hide"}
rm(list = ls())
library(MEGENA)
library(Seurat)
library(ggplot2)
library(cowplot)
library(Matrix)
```

Read in Seurat Object file.
```{r read Seurat object, eval = TRUE, echo = TRUE}
obj = readRDS("Output/SeuratObject.RDS")
```

# Setup some parameters.

```{r setup parameters, eval = TRUE, echo = TRUE}
n.cores <- 4;
doPar <- TRUE;
method = "spearman"
FDR.cutoff = 0.05
module.pval = 0.05
hub.pval = 0.05
```

As an example, we will construct MEGENA network for AD cells in cluster 5.
```{r get data, eval = TRUE, echo = TRUE}
clusterid = 5
obj.cls = obj[,(Idents(obj) == clusterid & obj@meta.data$pathologic.diagnosis.of.AD == 'YES')]

# remove unexpressed genes 
min.cells = ncol(obj.cls) * 0.05 # require at least 5% of cells expressed (~ 100 cells)
cnt = GetAssayData(obj.cls,slot = "counts")
nc = rowSums(cnt >= 1E-320,na.rm = TRUE)
ii = which(nc >= min.cells) # index of expressed genes

print(length(ii))

# remove noisy cells
min.genes = 100;
ng = colSums(cnt[ii,] > 1E-320,na.rm = TRUE)
jj = which(ng >= min.genes)

print(length(jj))

datExpr = GetAssayData(obj.cls,slot = "data")[ii,jj]

```

Prepare output directory.
```{r prepare output, eval = TRUE, echo = TRUE}
outdir = paste0('MEGENA/Cluster_', clusterid)
if(!file.exists(outdir)) dir.create(outdir, recursive = TRUE)
setwd(outdir)
```

Calculate pairwise correlation.
```{r calc corr, eval = TRUE, echo = TRUE}
ijw <- calculate.correlation(datExpr = as.matrix(datExpr), doPerm = 10, method = method, FDR.cutoff = FDR.cutoff)
```

Prep for parallelization prior to the PFN.
```{r recruite multi-cores, eval = TRUE, echo = TRUE}
if (doPar)
{
    cl = parallel::makeCluster(n.cores)
    doParallel::registerDoParallel(cl)
    # check how many workers are there
    cat(paste("number of cores to use:", foreach::getDoParWorkers(), "\n"))
}
```  
  
Calculate PFN.
```{r calc PFN, eval = TRUE, echo = TRUE}
el <- calculate.PFN(ijw[, 1:3], doPar = doPar, num.cores = n.cores)
write.table(el, file = "MEGENA_Network.txt", sep = "\t", row.names = FALSE, col.names = TRUE, quote = FALSE)
rm(ijw)
```

Do multiscale clustering.
```{r do clustering, eval = TRUE, echo = TRUE}
g <- graph.data.frame(el,directed = F)
MEGENA.output <- do.MEGENA(g,mod.pval = module.pval, hub.pval = hub.pval, remove.unsig = TRUE,
                          min.size = 20,#max.size = vcount(g)/2,
                          doPar = TRUE, num.cores = n.cores, n.perm = 100,
                          save.output = TRUE)
```

Save MEGENA analysis object in a file.
```{r save output, eval = TRUE, echo = TRUE}
save(MEGENA.output, file = "MEGENA_output.RData")
```

## Session info
```{r session info}
sessionInfo()
```
