---
title: 'ROSMAP snRNA-seq: Feature Selection to Clustering'
author: "Won-Min Song"
date: "November 18, 2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Run-down for snRNA-seq clustering analysis for ROSMAP data

This code includes 
The .RDS file for the pre-processed seurat object can be found at: https://www.synapse.org/#!Synapse:syn26477245. 


```{r load functions and packages}
##  Load necessary packages and functions into R
rm(list = ls())
library(Seurat)
library(ggplot2)
library(ggrepel)
library(scran)

print(getwd())
out.dir <- "clustering_results";dir.create(out.dir)

source("./scripts/R_functions/enrichment_functions.v2.R")
source("./scripts/R_functions/clustering_functions.v1.R")

```

## data loading

Now, since the data files is huge, load it from cloud

```{r data loading}
urlstr <- "https://www.dropbox.com/s/w0mzab5i3dwxwlp/SeuratObject.RDS?raw=1"
seu = readRDS(url(urlstr))
print(seu)
```

#### Get Feature selection
# use modelGeneVar function from scran
Here, we will demonstrate how to select genes with significant dispersions by using modelGeneVar() from scran workflow, or FindVariableFeatures() from Seurat

```{r model dispersion}
use.scran = FALSE # use scran workflow
use.seurat = TRUE # use seurat workflow

if (use.scran)
{
	gene.var = modelGeneVar(x = GetAssayData(seu),block = factor(seu@meta.data$individualID))
	gene.var =gene.var[order(gene.var$bio,decreasing = TRUE),]
	
	# get significant genes
	gene.sig.var = subset(gene.var,FDR < 0.05) # significant dispersion genes with FDR < 0.05

	# populate variable genes in seurat
	seu[["SCT"]]@var.features = rownames(gene.sig.var)

	library(ggplot2)
	library(ggrepel)
	pobj = ggplot() + geom_point(data = as.data.frame(gene.var[,1:6]),aes(x = tech,y = bio,colour = FDR < 0.05),alpha = 0.2) + 
	  scale_colour_manual(values = c("TRUE" = "red","FALSE" = "gray")) + 
	  geom_text_repel(data = subset(data.frame(gene.var[1:20,1:6],gene = rownames(gene.var)[1:20]),FDR < 0.05),aes(x = tech,y = bio,label = gene),size = 5) + # name top 20 genes for viz. 
	  labs(x = "Technical",y = "Biological") + 
	  guides(colour = guide_legend(title = "Biological Var.\nFDR < 0.05")) + 
	  theme_bw() + 
	  theme(axis.title = element_text(size = 20),axis.text = element_text(size = 17),
        legend.position = "bottom",legend.title = element_text(size = 20),
        legend.text = element_text(size = 17),legend.direction = "horizontal")

	png(file = paste0(out.dir,"/scran_variance.png"),res = 300,width = 1800,height = 2100)
	print(pobj)
	dev.off()
}

if (use.seurat)
{
	# Seurat's VST routine: note that nfeatures = 3000 is set. While Seurat recommends 2000, 3000 may be more liberal to include more variability in the data set. 
	seu = FindVariableFeatures(
	  object= seu,
	  selection.method = "vst",
	  nfeatures = 3000,
	  verbose = TRUE
	)
	# plot variable features with and without labels
	top10 <- head(VariableFeatures(seu), 10)
	plot1 <- VariableFeaturePlot(seu)
	plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
	png(file = paste0(out.dir,"/Seurat_variable_features.png"),res = 300,width = 3200,height = 2500)
	plot1 + plot2
	dev.off()
}

print(length(seu[["SCT"]]@var.features))

```

## PCA data
Now, let's perform PCA. This portion includes two pipelines for PC selection: getDenoisedPCs from scran, and observing explained variance curve for Seurat. 

```{r PCA}
use.scran = FALSE # use scran workflow
use.seurat = TRUE # use seurat workflow

### scran workflow to nominate PCs
if (use.scran)
{
	## get SCT normalized data from Seurat: But, this matrix can be any log-normalized data for scRNA-seq
	mat = GetAssayData(seu);
	mat = mat[match(rownames(gene.sig.var),rownames(mat)),]

	# in case you would want to denoise PCs
	pcs.denoised <- getDenoisedPCs(mat, technical=gene.sig.var) 
	
	# replace PCA components in Seurat object
	seu$pca@cell.embeddings = pcs.denoised$components[colnames(seu),]
	
	# run UMAP on updated PCs
	seu <- RunUMAP(seu, dims = 1:ncol(pcs.denoised$components), seed.use = 4867)
}


### seurat workflow to nominate PCs 
if (use.seurat)
{
	# Run PCA in seurat workflow
	seu <- ScaleData(seu, features = rownames(seu))
	seu = RunPCA(seu, features = VariableFeatures(object = seu))

	# Get the total variance:
	total_variance <- sum(matrixStats::rowVars(Seurat::GetAssayData(seu, assay = "SCT", slot = "scale.data")))
	eigValues =  (seu$pca@stdev)^2

	pca.pdata = data.frame(PC = 1:length(seu$pca@stdev),
	eigValues = (seu$pca@stdev)^2,  ## EigenValues
	varExplained = eigValues / total_variance * 100,
	totalExplained = cumsum(eigValues / total_variance * 100))

	pobj = ggplot(data = pca.pdata,aes(x = PC,y = varExplained)) + geom_point() + geom_line() + 
	geom_vline(xintercept = c(5,7,20),colour = c("red","blue","green")) + 
	theme_bw() + 
	labs(x = "PCs",y = "%. Variance Explained") + 
	theme(axis.title = element_text(size = 20),axis.text = element_text(size = 17))

	png(file = paste0(out.dir,"/variance_explained.png"),res = 300,width = 2400,height = 2100)
	print(pobj)
	dev.off()
	
	n.pcs = 20 # use 20 PCs after inspecting the plot
	seu <- RunUMAP(seu, dims = 1:n.pcs, seed.use = 4867)

}
```

## UMAP Embedding

```{r UMAP plotting}
# jointly visualize with other complex features
xy = Embeddings(seu,"umap")
pdata = cbind.data.frame(seu@meta.data,as.data.frame(xy[match(colnames(seu),rownames(xy)),]))
pdata$pathology.group = gsub("late","Late AD",gsub("early","Early AD",gsub("no","Healthy",gsub("-pathology$","",as.character(pdata$pathology)))))

# get cluster label coordinate
xy.o = do.call('rbind',lapply(split(1:nrow(xy),factor(seu@meta.data$seurat_clusters)),function(x,y) apply(rbind(y[x,]),2,function(q) median(q)),y = xy));colnames(xy.o) = c("x","y")
hobj = hclust(dist(xy.o),"complete")
cls.tick = hobj$label[hobj$order]

# first create comprehensive uMAP plot
colmap = c("Healthy" = "darkolivegreen1","Early AD" = "yellow","Late AD" = "coral")

pobj = ggplot() + 
  geom_point(data = pdata,aes(x = UMAP_1,y = UMAP_2,fill = pathology.group,colour = broad.cell.type),shape = 21,alpha = 0.4,size = 3,stroke = 1.5) + 
  geom_text(data = data.frame(cls.id = rownames(xy.o),as.data.frame(xy.o)),aes(x = x,y = y,label = cls.id),size = 7) + 
  scale_fill_manual(values = colmap) +
  guides(fill = guide_legend(title = "Diagnosis",ncol = 3),colour = guide_legend(title = "Cell Type",ncol = 3)) + 
  theme_bw() + 
  theme(axis.ticks = element_blank(),axis.title = element_blank(),axis.text = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.title = element_text(size = 20),legend.text = element_text(size = 18),
        legend.position = "bottom",legend.direction = "horizontal")

png(file = paste0(out.dir,"/umap.png"),res = 300,width = 3500,height = 2800)
print(pobj)
dev.off()
```

## Marker plotting on UMAP
```{r cell type markers}
# get marker expression plots
genes = c("Ex" = "NRGN","In" = "GAD1","Ast" = "AQP4","Ast" = "GFAP",
          "Mic" = "TYROBP","Mic" = "CX3CR1","Oli" = "PLP1","OPC" = "VCAN","End" = "FLT1")
dobj = DotPlot(seu, features = genes) + RotatedAxis() + scale_y_discrete(limits = cls.tick) + 
  labs(x = "Cell type marker",y = "Cluster") + 
  theme(axis.text.y = element_text(size = 16),axis.text.x = element_text(size = 14),
        legend.position = "bottom",legend.direction = "vertical")

png(file = paste0(out.dir,"/Celltype_Markers.Dotplot.png"),res = 300,width = 1700,height = 2700)
print(dobj)
dev.off()
```

## Composition analysis
```{r composition analysis}
## create proportion plot
# assign features 
mods = split(colnames(seu),Idents(seu));
vec = seu@meta.data$pathology.group;names(vec) = colnames(seu)
colmap = c("Healthy" = "darkolivegreen1","Early AD" = "yellow","Late AD" = "coral")

# check enrichments 
fet.res = perform.AllPairs.FET(geneSets1 = mods,geneSets2 = split(names(vec),factor(vec)),background = colnames(seu),adjust.FET.pvalue = T,or = 1,alternative = "greater",
                               do.multicore = F,n.cores = NULL)
fet.res = fet.res[order(fet.res$FET_pvalue),]
fet.res$corrected.FET.pvalue = p.adjust(fet.res$FET_pvalue,"BH")

require(reshape2)
pmat = acast(data = fet.res,formula = set1_Name ~ set2_Name,value.var = "corrected.FET.pvalue")

# get coordinates for marks
m = overlap_modules_wt_categ(mods,vec);
m = m[,match(rev(names(colmap)),colnames(m))]
dot.coord = t(apply(m,1,function(x) cumsum(x) - 0.5*x))

# make data for dot marks
dot.data = subset(fet.res,corrected.FET.pvalue < 0.05 & enrichment.foldchange > 1.2)
ij = cbind(match(as.character(dot.data$set1_Name),rownames(dot.coord)),match(as.character(dot.data$set2_Name),colnames(dot.coord)))
dot.data$y = apply(ij,1,function(x,m) m[x[1],x[2]],m = dot.coord)

# generate plot
probj = get_proportion_plot(modules = mods,vec = vec,cols = colmap)	
probj = probj + scale_fill_manual(values = colmap,limits = names(colmap)) + 
  geom_point(data = dot.data,aes(x = set1_Name,y = y,size = enrichment.foldchange),colour = "red",alpha = 0.5) + 
  guides(size = guide_legend(ncol = 1,title = "EFC"),fill = guide_legend(title = "Pathology",ncol = 1)) + 
  scale_x_discrete(limits = cls.tick) + scale_y_continuous(limits = c(0,1)) + 
  coord_flip() + 
  theme(axis.text = element_text(size = 19),
        legend.text = element_text(size = 18),legend.title = element_text(size = 20))


png(file = "Pathology_Proportion.png",res = 300,width = 1340,height = 2800)
print(probj)
dev.off()
```


## Session info
```{r session info}
sessionInfo()
```