---
title: 'ROSMAP snRNA-seq: Copy Number Variation Detection Session'
author: "Chen Ming"
date: "Nov 30th, 2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 
This tutorial show how to run InferCNV to detect CNVs in Cluster6 of snRNA-seq of the ROSMAP cohort.  
The original InferCNV Tutorial: https://github.com/broadinstitute/inferCNV/wiki/Running-InferCNV  
Before running codes below:  
module load jags  
R version: R/4.0.2  

Load necessary packages and functions into R  

```{r load functions and packages}
rm(list = ls())
library(infercnv)
library(Seurat)
library(dplyr)

outdir <- "CNV_results"
if(!file.exists(outdir)) dir.create(outdir, recursive = TRUE)

```

## data loading

Now, since the data files is huge, load it from cloud.
A precomputed version of the SeuratObject file is available at https://www.synapse.org/#!Synapse:syn26477245
Please download the precomputed SeuratObject first.

```{r data loading}
seurat_obj=readRDS("SeuratObject.RDS")
print(seurat_obj)
```

# Subset Cluster 6:
Here, we extract cells of Cluster 6

```
mycluster=subset(x = seurat_obj, idents = 6)
```
# Preparation before creating the infercnv object
# Raw count matrix:
Next, we need to prepare several variables before constructing the InferCNV object
Extract raw count matrix:

```
raw_counts_matrix=mycluster@assays$RNA@counts
```

# Cell annotation file:
Prepare cell annotation file:

```
cellnames=row.names(mycluster@meta.data)
cellclusters=paste(mycluster@meta.data$SeuratCelltype,mycluster@meta.data$pathology.group,sep="_")
mycellAnnotation=t(t(cellclusters))
rownames(mycellAnnotation)=cellnames

```

# Construct InferCNV object
Now we construct the inferCNV object  
The "my_gen_pos.GRCh38.txt" is prepared gene location file of GRCh38.  

```
infercnv_obj = CreateInfercnvObject(raw_counts_matrix=raw_counts_matrix,annotations_file=mycellAnnotation,delim="\t",gene_order_file="my_gen_pos.GRCh38.txt",ref_group_names=unique(cellclusters)[grep("no-pathology",unique(cellclusters))])					
```
	
# CNV detection
Now we run InferCNV to detect CNV:

```
infercnv_obj = infercnv::run(infercnv_obj,cutoff=0.1, out_dir=outdir cluster_by_groups=TRUE, denoise=TRUE, HMM=TRUE)
```

#
The CNV regions are stored in:HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_regions.dat and HMM_CNV_predictions.HMMi6.hmm_mode-samples.Pnorm_0.5.pred_cnv_genes.dat