---
title: 'ROSMAP snRNA-seq: Bayesian network analysis'
author: "Minghui Wang"
date: "November 29, 2021"
output: html_document
---


#### Step 1. Prepare data for Bayesian network analysis using RIMBANet
```{r setup environment, eval = TRUE, echo = TRUE, results = "hide"}
rm(list = ls())
library(Seurat)
library(Matrix)
```

Read in Seurat Object file.
```{r read Seurat object, eval = TRUE, echo = TRUE}
obj = readRDS("Output/SeuratObject.RDS")
```

First, extract the cluster-specific expression data. As an example, we will use data from cells in cluster 5.
```{r get data, eval = TRUE, echo = TRUE}
clusterid = 5
#extract data and filter lowly expressed genes
expr_dat = GetAssayData(obj[, Idents(obj) == clusterid], slot = 'data')
expr_dat = expr_dat[rowSums(expr_dat > 0) >= 0.2 * ncol(expr_dat), ]
expr_dat = expr_dat[apply(expr_dat, 1, function(x) length(unique(x))) > 3, ]
```

Prepare output directory.
```{r prepare output, eval = TRUE, echo = TRUE}
outdir = paste0('BN/Cluster_', clusterid, '/')
if(!file.exists(outdir)) dir.create(outdir, recursive = TRUE)
setwd(outdir)
```

Define a discretization function. Here we will use k-means clustering (k = 3) to discretize the expression data.
```{r k-means discretization, eval = TRUE, echo = TRUE}
discretizeK = function(x, k = 3){
	y = kmeans(x, k)
	order(y$centers)[y$cluster] - 1
}
expr_dat_3 <- t(apply(expr_dat, 1, discretizeK))
```

Write discretized data into file `data.discretized.txt`.
```{r save data, eval = TRUE, echo = TRUE}
write.table(expr_dat_3, file = paste0(outdir, 'data.discretized.txt'), col.names = FALSE, row.names = TRUE, quote = FALSE, sep = "\t")
```

#### Step 2. Construct BN by using RIMBANet
Now run the RIMBANet analysis with `data.discretized.txt` by following a pipeline at https://github.com/mw201608/BayesianNetwork

## Session info
```{r session info}
sessionInfo()
```
